---
title: "Introduction to the {isoplants} package"
author: "David Basler"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{isoplants: Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  eval = TRUE
)
```

The {isoplants} package allows to calculate isotopic composition of plant tissue in respone to environmental factors. 
This vignette shows a basic example to usse the {isoplants} package:

- run a minimum worked example using default parameters
- replace default parameters
- include uncertainty estimates for specific parameters

## Minimum worked example

You can use the models with the default parameter settings using the `get_default_parameters()` function and `plant18O_model()`. Basic information about the inputparmeters can be displayed by calling `get_parameter_definition()`.


```{r, basic}

library(magrittr)
library(isoplants)

# Get all default parameters (the default parameters will use include parameters for the peclet model)
df_parameter<-get_default_parameters()
result<-plant18O_model(df_parameter)
result %>% knitr::kable()


# The output can also be limited to a specific variable
plant18O_model(df_parameter,output = 'd18O_c')
```

## Parameter Checking

The package includes the `check_parameters()` function which provides 
(1) a list of selected model options, 
(2) basic checking of provided parameters and 
(3), a list of any errors found.


```{r, check}

check<-check_parameters(df_parameter) # Run a check on the specified parameters

check$model_options %>% knitr::kable()
check$model_parameters %>% knitr::kable()
check$errors %>% knitr::kable() # should be empty

```

## Replace parameters

The input parameter data.frame object can have multiple rows. Here we calculate d18O_lt for different values of RH.

```{r, cpar}
# Run model on default parameters, manipulating one parameter (RH)
RH<-seq(10,90,10)
# PArameters can be set using the set_parameter function
df_parameter<-get_default_parameters()
df_parameter<-set_parameters(data.frame(RH),df_parameter)
# OR by assigning them directly to a parametertable with the appropriate number of rows
df_parameter<-get_default_parameters(n=length(RH), mode='peclet')
df_parameter$RH<-RH
#run model
result<-plant18O_model(df_parameter)
plot(df_parameter$RH,result$d18O_bulk,xlab='RH',ylab='d18O_lt',pch=19,las=1)
```

## Parameter uncertainty

The included fuction `get_randomized_parameters()` samples n parametersets from parameter specific distributions in order to estimate the effect of uncertainty of input parameters. `set_parameters()` combines fixed parameters with a set of varying parameters.

```{r, unc}
# Run model on default parameters with uncertainty on multiple parameter

# Get default parameters
df_parameter<-get_default_parameters(mode='peclet')
# Defines ranges and distributions for parameters (see ?get_randomized_parameters for detailed information on the imput format and further options)
rnd_par <- data.frame(name= c('RH',   'd18O_sw', 'px'   ),
                      pdist=c('norm', 'norm'   , 'unif' ),
                      pdm=  c(70,      -10     , 0.4    ),
                      pdv=  c(20,       5      , 0.2    ),
                      stringsAsFactors = FALSE
                      )

n<-1000 # Number of samples

df_parameter <- set_parameters(get_randomized_parameters(n,rnd_par) , df_parameter)
result<-plant18O_model(df_parameter)

```
```{r, fig.show='hold'}
hist(result$d18O_pt,main = 'd18O_pt',xlab='d18O plant tissue')
boxplot(result$d18O_pt,main = 'd18O_pt')

```


