---
title: "Isotope model sensitivity analysis with the {isoplants} package"
author: "David Basler"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{isoplants: Sensitivity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  eval = TRUE
)
```

The models in the {isoplants} package are include various environmental and leaf specific input parameters. The package includes functions to analyze the sensitivity to changes in the input variables
This vignette shows some basic example to do some seinsitivity analysis with {isoplants}:


## Direct sensitivity analysis

For example, we can look at how d18O of cellulose is affected by differnt pex values under a set of differnet RH conditions. 

```{r sa}
library(isoplants)

pex<-seq(0.1,0.9,0.01)
lf18Opar<-get_default_parameters(n=length(pex))

lf18Opar$pex<-pex
result<-plant18O_model(lf18Opar)

```

```{r fig.width=6,fig.height=4}

plot(lf18Opar$pex,result$d18O_c,ylim=c(0,25),xlim=c(0,1),type='l',las=1,xlab='pex',ylab='d18O cellulose')
rhs<-seq(10,90,10)
col<-rainbow(length(rhs))
for (i in 1:length(rhs)){
  lf18Opar$RH<-rhs[i]
  result<-plant18O_model(lf18Opar)
  lines(lf18Opar$pex,result$d18O_c,col=col[i])
}
legend('bottomleft',as.character(rhs),col=col,lty=1,title = 'RH (%)',bty='n')

```


## Sobol incices

Sobol incices are a Global form of sensitivity indices whereby the variance of the output of the model is decomposed into fractions which can be attributed to individual input parameters or from the interaction between parameters. Here we analyze the sensitivity of d18O cellulose to changes in the input parameters.

```{r sobol}
# Sensitivity analysis (with sobol indices)

library('sensitivity')
library('boot')
library('ggplot2')

# Define parameter ranges (min-max)
rnd_par <- data.frame(
  name  = c("Tair","DTleaf","RH",  "P", "d18O_sw","D18O_wv","rb", "Lm", "gsref", "px", "pex"),
  pdist = 'minmax',
  pdm   = c(     5,     -5,  30,   90.0,       -30,     -20, 0.4, 0.001,    0.01 ,  0.8,  0.4),
  pdv    = c(    30,      10,  95, 102.0 ,       0,       0,   5,    0.5,     0.6 ,  0.8,  0.4),
  stringsAsFactors = FALSE
)
#Initialize two random parametersets
n<-10000
X1 <- get_randomized_parameters(n,rnd_par)
X2 <- get_randomized_parameters(n,rnd_par)

x <- sobol2007(model =plant18O_model, X1 = X1, X2 = X2, nboot = 1000, output = "d18O_c")
```
```{r, fig.width=6,fig.height=4}
ggplot(x)

```

### Fixing some parameters
The `plant18O_model()` model allows to pass some parameters with the `addpar` argument. These will be internally joined with the main parameters, but allow for the exclusion of these parameters in certain kinds of analysis. 
Here we analyze the sensitivity of d18O leafwater to changes in the input parameters (excluding variation in soil water, and the parameters for the steps after the calculation of leaf water).

```{r sobolfix}
# Calculate Sobol indices with a set of fixed parameters (set to default values)
fixpar <- get_default_parameters()[,c('d18O_sw','Lm','px','pex','ecp')]
n<-10000
# Using the previously defined parameter ranges (rnd_par), excluding the fixed parameters
X1 <- get_randomized_parameters(n,rnd_par)[,!names(fixpar) %in% names (rnd_par)]
X2 <- get_randomized_parameters(n,rnd_par)[,!names(fixpar) %in% names (rnd_par)]
X1<-X1[,!names(X1) %in% names (fixpar)]
X2<-X2[,!names(X2) %in% names (fixpar)]

x <- sobol2007(model =plant18O_model, X1 = X1, X2 = X2, nboot = 1000, output = "d18O_lw",addpar=fixpar)
```
```{r, fig.width=6,fig.height=4}
ggplot(x)

```


### Parmeter gradients

In order to analyze the importance of multiple input parameters across a gradient of a certain parameter, we can use the `scan_sensitivity()` function

```{r sobolscan}
n<-10000
X1 <- get_randomized_parameters(n,rnd_par)
X2 <- get_randomized_parameters(n,rnd_par)
fixpar <- get_default_parameters()[,c('d18O_sw','P','px','pex')]
scanpar<-"Tair"
ts<-seq(0,30,2)
ss<-scan_sensitivity(X1,X2,fixpar,ts,scanpar,outpar="D18O_lw")
```

```{r,fig.width=6,fig.height=4}
plot_sensitivity(ss$ds,scanpar,'normalized first-order indices','D18O_lw',parcol = 1)


```
```{r,fig.width=6,fig.height=4}
plot_sensitivity(ss$dt,scanpar,'normalized total indices','',parcol = 1,legend = 2)
```
