---
title: "Introduction to the {isoplants} package"
author: "David Basler"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{isoplants: Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  eval = TRUE
)
```

The {isoplants} package allows to calculate isotopic composition of plant tissue in respone to environmental factors. 
This vignette shows a basic example to use the {isoplants} package:

- run a minimum worked example using default parameters
- replace default parameters
- include uncertainty estimates for specific parameters

## Minimum worked example

You can use the models with the default parameter settings using the `get_default_parameters()` function and `plant18O_model()` or `plant2H_model()`. Basic information about the interpreters can be displayed by calling `get_parameter_definition()`.


```{r, basic}

library(magrittr)
library(isoplants)

# Get all default parameters (the default parameters will use include parameters for the Peclet model)
df_parameter <- get_default_parameters() # for oxygen [default]
result <- plant18O_model(df_parameter)
result %>% knitr::kable()

# The output can also be limited to a specific variable
plant18O_model(df_parameter,output = 'd18O_c')

# Similarly, this can also be applied for the hydrogen model
df_parameter <- get_default_parameters(element="H")
result <- plant2H_model(df_parameter)
result %>% knitr::kable()

```

## Parameter Checking

The package includes the `check_parameters()` function which provides 
(1) a list of selected model options, 
(2) basic checking of provided parameters and, 
(3) a list of any errors found.


```{r, check}
df_parameter <- get_default_parameters(element="O")
check<-check_parameters(df_parameter) # Run a check on the specified parameters
check$model_options %>% knitr::kable()
check$model_parameters %>% knitr::kable()
check$errors %>% knitr::kable() # should be empty

```

## Replace parameters

The input parameter data.frame object can have multiple rows. Here we calculate d18O_lt for different values of RH.

```{r, cpar}
# Run model on default parameters, manipulating one parameter (RH)
RH<-seq(10,90,10)
# PArameters can be set using the set_parameter function
df_parameter<-get_default_parameters()
df_parameter<-set_parameters(data.frame(RH),df_parameter)
# OR by assigning them directly to a parameter table with the appropriate number of rows
df_parameter<-get_default_parameters(n=length(RH), mode='peclet')
df_parameter$RH<-RH
#run model
result<-plant18O_model(df_parameter)
plot(df_parameter$RH,result$d18O_pt,xlab='RH',ylab='d18O_lt',pch=19,las=1)
```

## Parameter uncertainty

The included fuction `get_randomized_parameters()` samples n parametersets from parameter specific distributions in order to estimate the effect of uncertainty of input parameters. `set_parameters()` combines fixed parameters with a set of varying parameters.

```{r, unc}
# Run model on default parameters with uncertainty on multiple parameter

# Get default parameters
df_parameter<-get_default_parameters(mode='peclet')
# Defines ranges and distributions for parameters (see ?get_randomized_parameters for detailed information on the imput format and further options)
rnd_par <- data.frame(name= c('RH',   'd18O_sw', 'pex'   ),
                      pdist=c('norm', 'norm'   , 'unif' ),
                      pdm=  c(70,      -10     , 0.4    ),
                      pdv=  c(20,       5      , 0.2    ),
                      stringsAsFactors = FALSE
                      )

n<-1000 # Number of samples

df_parameter <- set_parameters(get_randomized_parameters(n,rnd_par) , df_parameter)
result<-plant18O_model(df_parameter)

```
```{r, fig.show='hold'}
hist(result$d18O_pt,main = 'd18O_pt',xlab='d18O plant tissue',las=1)
boxplot(result$d18O_pt,main = 'd18O_pt',las=1)

```


## Optimize a set of parameters

Quite often, a set of parameters need to be fitted to some training data in order to produce a operational model.{Isoplants} allows for very simple parameter optimization when measured data is available.

```{r, pfit}

# prepare dataset with some randomized sample data
n<-40
observed_data <- get_default_parameters(n)
observed_data$Tair <- rnorm(n,20,4)
observed_data$RH   <- rnorm(n,70,10)
observed_data$Lm   <- 0.06  # The fitting procedure should produce something close to this value
observed_data$pex   <- 0.5   # The fitting procedure should produce something close to this value
# create some 'measured' vaules are added to the data.frame as 'd18O_cellulose'
# (ensure not to use a parameter name)
noise<-(runif(n)-0.5)/10
observed_data$d18O_cellulose <- plant18O_model(observed_data,output='d18O_c') + noise
# Always ensure that there are no NA values in the observed values
observed_data<-observed_data[!is.na(observed_data$d18O_cellulose),]


# Fit the parameters "Lm" and "px" (this might take a while)
parfit<-fit_plant18O(c("Lm","pex"), observed_data, obsvalue = "d18O_cellulose", modvalue ="d18O_c")

print (parfit$par)
# Run model on best parameters
run_par<-get_run_parameters(observed_data,parfit$par)
modelout<-plant18O_model(run_par)
result<-cbind(observed_data,modelout)

```
```{r, fig.show='hold'}

# Plot predicted vs observed
plot(d18O_cellulose~d18O_c,data=result,xlab='d18Oc Predicted',ylab='d18Oc Observed',las=1,pch=19,col='red')
segments(0,0,100,100,lty=2)
abline(lm(d18O_cellulose~d18O_c,data=result),col='red')

```




## Leaf temperature model

The `isoplants` package is designed to work with the leaf temperature modelling package `tealeaves` (Muir 2019) to estimate leaf temperature in various environmental conditions or integrate the feedback effect of stomatal conductance on leaf temperature into the isoplants model.


```{r, ltemp}

# Get a isoplants parameterset, including default values for radiation windspeed and leaf size
df_parameter<-get_default_parameters(tealeaves = TRUE) 

# Set swrad to irradiance at 800 w/m2
df_parameter$swrad<-800

#set a range of values for stomatal conductance
df_parameter<-set_parameters(data.frame(gs=seq(0.04,0.5,0.01)),df_parameter) 
#update the  parameter set with Tleaf and rb as estimated from the tealeaves model
df_parameter<-calculate_Tleaf(df_parameter) 
# run the 18O isotope model
result_with_Tleaf<-plant18O_model(df_parameter)

# Run again setting Tleaf=Tair to show the effect of integrating Tleaf
df_parameter_without_Tleaf<-df_parameter
df_parameter_without_Tleaf$Tleaf<-df_parameter_without_Tleaf$Tair
result_without_Tleaf<-plant18O_model(df_parameter_without_Tleaf)


```
```{r, fig.show='hold'}

plot(df_parameter$Tleaf~df_parameter$gs,las=1,xlab='gs (mol m-2 s-1)',ylab='Leaf tempertaure (°C)',type='l')
text(0.5,19.7,"Air temperature 20°C",adj=c(1,1),cex=0.8)

plot(result_with_Tleaf$d18O_pt~df_parameter$gs,las=1,xlab='gs (mol m-2 s-1)',ylab='d18O plant tissue (‰)',type='l',lwd=2,col="#00AFBB")
lines(result_without_Tleaf$d18O_pt~df_parameter_without_Tleaf$gs,lty=2,lwd=2,col="#FC4E07")
legend('bottomleft',lty=c(1,2),lwd=2,c('incl.leaf temperature effect','excl. leaf temperature effect'),col=c(col="#00AFBB","#FC4E07"),bty='n',cex=0.8)

```


## Piso.AI API interaction

One of the most important parameter for the isotopic composition of plant tissue is the isotopic composition of source water (d18O_sw). Often this is approximated by using site-specific data on the isotopic composition of precipitation. The {isoplants} package interacts with the Piso.AI API to these data. [Piso.AI](https://isotope.bot.unibas.ch/PisoAI/) is tool for predicting monthly time series of oxygen and hydrogen isotope values of precipitation that uses a machine learning model trained on geographic and climate data. The Piso.AI v1.0 data set covers most of Europe and covers the time span from 1950-01-01 to 2019-12-31. 
Below is an example of the basic data requests from the Piso.AI api using the tools provided in {isoplants}:

```{r, Piso.AI, eval=FALSE}
# Simple request for Piso.AI data of two locations and the full time span of Piso.AI
location<-data.frame(site= c("TEST1","TEST2") ,latitude=c(45.7,46.5),longitude=c(7.6,8.0))
pisoai_data <- get_pisoai_data(location)

# Request for Piso.AI data of two locations and specific timespan
pisoai_data <- get_pisoai_data(location, years=c(1990,2000),months=c(4,6))

# Request for Piso.AI data of two locations individual years
location<-data.frame(site= c("TEST1","TEST2") ,latitude=c(45.7,46.5),longitude=c(7.6,8.0),years=c(1979,2010),stringsAsFactors = FALSE)
pisoai_data <- get_pisoai_data(location)

# Similar request, additionally storing data locally
pisoai_data <- get_pisoai_data(location,storelocal='~/PisoAI_data')
# read back local data
pisoai_data <- pisoai_readlocal('~/PisoAI_data')
```
